The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-11 rational.tex
sec fractional types

The section follows the outline of the previous one, first presenting the new
syntax, and discussing the graphical language and the properites implied by
the underlying categorical structure.

Multiplicative duality is added to {{langRev}} by adding types
{{1/b}}, values {{1/v}} and isomorphisms {{eta*}} and {{eps*}} with
the following type judgements.

@@

\vspace{-15pt}
\begin{multicols}{2}
%subcode{opsem} include main
% eta* &: 1 <-> (1/b) * b :& eps*

@@

As with additive duals, we represent {{eta*}} and {{eps*}} using
U-shaped connectors. In the general case the wires for {{1}} will be
dropped as it can freely introduced or eliminated using {{uniti}} and
{{unite}}.

@@

The usual interpretation of {{b1*b2}} we have both a value of type
{{b1}} and a value of type {{b2}}. This interpretation is maintained
in the presence of fractionals. Hence an {{eta* : 1 <-> (1/b) * b}} is
to be viewed as a fission point for a value of type {{b}} and its
multiplicative inverse {{1/b}}. This naturally raises the question --
which value of type {{b}} is produced? The isomorphism {{eta*}} does
not produce any specific value of type {{b}}, instead it produces a
place holder for a value such that the value of type {{1/b}} is
always the dual of the chosen one.

@@

The constructions for functions, involution, and trace are identical,
replacing {{+}} by {{*}} mutatis mutandis.

@@

>>>>

%% The section follows the outline of the previous one, first presenting
%% the new syntax, and discussing the graphical language and the
%% properites implied by the underlying categorical structure.

Multiplicative duality is added to {{langRev}} by adding types
{{1/b}}, values {{1/v}} and isomorphisms {{eta*}} and {{eps*}}.

@@

These have type judgements, similar to those of their additive
couterparts.
\begin{multicols}{2}
%subcode{opsem} include main
% eta* &: 1 <-> (1/b) * b :& eps*

@@

As with additive duals, we represent {{eta*}} and {{eps*}} using
U-shaped connectors. In later diagrams, the wires for {{1}} will be
dropped as they can freely introduced or eliminated using {{uniti}}
and {{unite}}.

@@

The usual interpretation of {{b1*b2}} we have both a value of type
{{b1}} and a value of type {{b2}}. This interpretation is maintained
in the presence of fractionals. Hence an {{eta* : 1 <-> (1/b) * b}} is
to be viewed as a fission point for a value of type {{b}} and its
multiplicative inverse {{1/b}}. This naturally raises the question --
which value of type {{b}} is produced? The isomorphism {{eta*}} does
not produce any specific value of type {{b}}, instead it produces a
placeholder for an unspecified value and dually, {{1/b}} is inhabited
by the placeholder that indicates the absence of the unspecified
value.

@@

@@

The constructions for functions, involution, and trace are identical,
replacing {{+}} by {{*}}, \textit{mutatis mutandis}.
